ERNIE: Accessible Support for Research Assessment

### 2017-12-26 LaTeX/Paper2/paper2.tex
Paper updates

Using data mining of publicly available administrative, regulatory, and scientific records and network analysis, Williams and colleagues defined relationships between scientific discoveries and major advances in medicine such as new drugs~\cite{Williams2015}. Extending this work, we have previously documented the scientific collaborations that extend across networks underlying the development of five independently developed therapeutics for cancer~\cite{Keserci2017}. Analysis of such networks supports new insights into collaboration, knowledge diffusion, and recursive learning \textit{(supra vide)}. The framework we developed to mine data and integrate it into networks for analysis has the benefit of retaining expert inputis flexible but can be further automated, adapted to a range of subjects beyond drug development, and used to support individual researchers, groups, and research organizations engaged in research assessments. In pursuit of these objectives we have developed Enhanced Research Network Informatics Environment (ERNIE), a knowledge platform that integrates publicly available data as well as data from commercial sources. Since cloud computing has evolved to the point where scalable computing services and storage are accessible to a broad user population, ERNIE was designed to exist as a cloud resource.

@@

\emph{Infrastructure} ERNIE exists as two Centos 7.4 virtual machines in the Microsoft Azure cloud with 10 terabytes of attached storage. One virtual machine,  standard D8s v3 (8 vcpu, 32 Gb), houses data in a PostgreSQL 9.6 database. A second virtual machine, a standard DS4 (8 vcpu, 28 Gb), has Apache Solr 7.x installed and is used to enable fast text based searches of the data in ERNIE. Automated processes are managed through the Jenkins Continuous Integration server and use custom ETL processes in Python to load and refresh data into the ERNIE schema. Secure access is managed through multiple factor authentication. Ssh tunnels are used to facilitate communications between virtual machines. Specifications, deployment scripts, and ETL scripts are available on a Github site~\cite{GithubERNIE2017}. \emph{[Need some text from DK and Avon!]}

\emph{Data sources} Data in ERNIE are derived from publicly available and commercial sources. Publicly available data are copied from the National Clinical Trials Database, the FDA Orange and Purple Books,  the United States Patent Office, and NIH ExPORTER. Leased data are acquired from Clarivate Analytics and consist of the Web of Science Core Collection and the Derwent Patent Citation Index. The data in ERNIE are stored in a relational schema that is also documented in Github.

\emph{Data mining and digitization} A set of source documents relevant to the question being asked is assembled and mined for citations to scientific literature using manual or semi-manual methods. Manual extraction involved searching for text strings that represent references of interest and matching those to unique identifiers in PubMed and/or Web of Science. As an example, the medical review document relevant to approval of ipilimumab contained X references. Y of them were matched to PubMed identifiers using the PubMed either through its web interface or the E-utils API. This process was modified to take advantage of the Solr search engine.

>>>>

Using data mining of publicly available administrative, regulatory, and scientific records and network analysis, Williams and colleagues defined relationships between scientific discoveries and major advances in medicine such as new drugs~\cite{Williams2015}. Extending this work, we have previously documented the scientific collaborations that extend across networks underlying the development of five independently developed therapeutics for cancer~\cite{Keserci2017}. Analysis of such networks supports new insights into collaboration, knowledge diffusion, and recursive learning \textit{(supra vide)}. The framework we developed to mine data and integrate it into networks for analysis has the benefit of retaining expert input is flexible but can be further automated, adapted to a range of subjects beyond drug development, and used to support individual researchers, groups, and research organizations engaged in research assessments. In pursuit of these objectives we have developed Enhanced Research Network Informatics Environment (ERNIE), a knowledge platform that integrates publicly available data as well as data from commercial sources. Since cloud computing has evolved to the point where scalable computing services and storage are accessible to a broad user population, ERNIE was designed to exist as a cloud resource.

@@

\emph{Open Source} All ERNIE code is open-source under the MIT license and available on GitHub ~\cite{GithubERNIE2017}.

\emph{Infrastructure} ERNIE runs on two CentOS 7.4 Linux servers in the Microsoft Azure cloud with 10 terabytes of
attached storage.
One server, Azure standard D8s v3 (8 VCPUs, 32 Gb RAM), hosts data in a PostgreSQL 9.6 relational database, and
additionally acts as a central hub for data related processes.
A second server, an Azure standard DS4 (8 VCPUs, 28 Gb), runs Apache Solr 7.x search server, which is used to to create
and store indexes of ERNIE data to facilitate fast free-text user searches.
Access to ERNIE servers is secured through SSH tunneling and two-factor authentication, with a unique time-based code
acting as one of the factors.

Automated ERNIE processes are managed by the Jenkins Continuous Integration server.
Custom Extract-Transform-Load (ETL) scripts are developed in Bash, Python, Java and SQL.
Jenkins is additionally used to deploy scripts dedicated to data transformation and perform other
calculations.
In that regard, Jenkins is used to direct construction of Solr indexes as new data is introduced into
ERNIE, as well as the construction of citation networks for specified topics.
The Solr indexes can handle varying forms of Lucene-based search and full-text search, and may quickly return
informative, ranked, and scored query results so that a user may be able to map data across sources.
The citation networks are sub-networks (based largely in the Web of Science) that are generated based upon user input
and subsequently analyzed.

\emph{Data sources} Data in ERNIE are extracted from publicly available and commercial sources. Publicly available data are extracted from the National Clinical Trials Database, the FDA Orange and Purple Books, the United States Patent Office, and NIH ExPORTER. Leased data are acquired from Clarivate Analytics and consist of the Web of Science Core Collection and the Derwent Patent Citation Index. The data in ERNIE are stored in a relational schema that is documented. Documentation is available on GitHub.

\emph{Data architecture and governance} ERNIE is aiming to provide data of best possible quality to
its analysts and users.
ERNIE data is extracted from several upstream sources of varying quality hence the need to cleanse
data.
ERNIE's approach to cleansing is to prefer data quality and normalized data over data completeness.
The following are key aspects of ERNIE's data architecture and governance:
\begin{itemize}
  \item De-duplication.
De-duplication is a process of eliminating logically duplicate records, that is records with
identical primary or unique keys.
In ERNIE duplicate records from such a group are deleted while leaving in place a single
representative record (preferring a record which is updated last).
An alternative de-duplication strategy could be to aggregate differing non-key fields from logical
duplicates into lists.
  \item Normalization.
After data is de-duplicated, the next step in data normalization is to establish foreign key
relationships between children and parent entities and eliminate orphans: child records with
non-existing parents.
In ERNIE orphan records are deleted.
An alternative normalization strategy could be to segregate orphans logically or physically while
creating parent stub records.
  \item Schema Permanence.
In order to improve availability of data and simplify data schema management, ERNIE ETL scripts are
designed to avoid recreating of main entities, for example, by using TRUNCATE TABLE instead of table re-creation.
  \item Data Stewardship.
ERNIE data stewards are people responsible for quality of a particular data subject area.
Data stewards get notified when a data issue is detected and drive the workflows necessary for data
issue remediation.
They are also typically responsible for communicating changes in data structures.
\end{itemize}

\emph{Data mining and digitization} A set of source documents relevant to the question being asked is assembled and mined for citations to scientific literature using manual or semi-automated methods. Manual extraction involved searching for text strings that represent references of interest and matching those to unique identifiers in PubMed and/or Web of Science. As an example, the medical review document relevant to approval of ipilimumab contained X references. Y of them were matched to PubMed identifiers using the PubMed either through its web interface or the E-utils API. ERNIE is designed to support semi-automated case studies.
